---
title: "Tidy Data - Assignment 6 (IS607)"
author: "Arun Kumar"
date: "October 4, 2015"
output: 
    html_document:
      toc: yes
---

###Required libraries  
```{r}
library(knitr); #For using kable functions for tables  
library(tidyr); #For tidying data
library(dplyr); #For easy data scrubbing and manipulation
```
###The raw data  
####Data download from Github  
The data is uploaded into my Github repository. Downloading the data to an R object.  
```{r}
flights_data_master <- read.table(file = "https://raw.githubusercontent.com/arunk13/MSDA-Assignments/master/IS607Fall2015/Assignment6/FlightStatus.csv", stringsAsFactors = FALSE, sep = ",", header = TRUE, check.names = FALSE);
flights_data <- flights_data_master; #Working copy
```

Below is the raw data :

```{r}
kable(flights_data, align = c('c'));

```

####Variables in the dataset  
I find the below variables in my dataset :  
a. Airlines  
b. Arrival_status  
c. City  

###Data Scrubbing  
####1. Converting wide to long format  
For easier analysis, I will like to convert my dataset from a wide format to a long format. :  

```{r}
flights_data_long <- flights_data %>% 
                  gather("City", "Stats", 3:ncol(flights_data));
kable(head(flights_data_long), align = c('c'));
```

####2. Data by delay types  
As we have to do analysis on arrivals, I think that it will be a good idea to spread the arrival and delay times  separately.  
```{r}
flights_data_long <- flights_data_long %>% 
                       spread(Arrival_Status, Stats) %>% 
                         group_by(Airlines, City);
kable(flights_data_long, align = c('c'));
```

###Data Analysis  
####Arrival delays by airlines  
```{r}
flights_data_final <- flights_data_long %>% 
                          group_by(Airlines) %>% 
                              summarise(delayed_arrivals = sum(Delayed), on_time_arrivals = sum(`On Time`),`delayed_arrivals(%)` = round((delayed_arrivals /(delayed_arrivals+on_time_arrivals))*100, digits = 2)) %>% select(Airlines, `delayed_arrivals(%)`);

kable(flights_data_final, caption = "Arrival delay comparison by airports", align = c('c'));
```

###Conclusion

####The Alaska airlines have more arrival delays than the AM West airlines. 
