---
title: 'IS607: Week 4 Assignment'
author: "Arun Kumar"
date: "September 19, 2015"
output: html_document
---


       
```{r echo=FALSE}
library(stringr);
```     
      
####4. Describe the types of strings that conform to the following regular expressions and construct an example that is matched that is matched by the regular expression.      
a. [0-9]+\\$  
Answer:  Finds a number having digits 0-9 of any length and ends with $.
```{r}
question.one <- "My number is 1236534512 but i can also be reached through 2124496571$";  
unlist(str_extract_all(question.one, "[0-9]+\\$"));
```
b. \\b[a-z]{1,4}\\b  
Answer: Finds words made up of lower case alphabets (a to z) of length 1 to maximum lenth of 4.
```{r}
unlist(str_extract_all(question.one, "\\b[a-z]{1,4}\\b"));
```
c. .*?\\.txt$
Answer: Finds any match which ends with ".txt". As its greedy and has ".*", it will identify the whole sentence if the sentence ends with .txt.
```{r}
my.file <- "My name 1 is myPersonalDetails.txt";
unlist(str_extract_all(my.file, ".*?\\.txt$"));
```
d. \\d{2}/\\d{2}/\\d{4}
Answer: Finds any part of the sentence which has number in format - two numbers/two numbers/four numbers
```{r}
random.number <- "This is a random abx333/44/333322";
unlist(str_extract_all(random.number, "\\d{2}/\\d{2}/\\d{4}"))
```
e. <(.+?)>.+?</\\1>
Answer: Find tags wich is a format "<something>content</something>"
```{r}
random.tag <- "<html>Hellow from Arun</html>";
unlist(str_extract_all(random.tag, "<(.+?)>.+?</\\1>"));
```
####5. Rewriting the same expression [0-9]+\\$     
```{r}
unlist(str_extract_all(question.one, "\\d*\\$"));
```
####6.  
```{r}
my.email <- "chunkylover53[at]aol[dot]com";
```
a. Transforming the string to standard mail format.
```{r}
my.email <- str_replace(my.email, "\\[at\\]", "@");
my.email <- str_replace(my.email, "\\[dot\\]", ".");
```
  
The transformed email address : `r my.email`     
b. What's wrong in using [[:digit:]] ?
```{r}
unlist(str_extract_all(my.email, "[[:digit:]]"));
```
#####The expression [[:digit:]] will only identify the first occurence.  
#####To get the complete number, we need a greedy expression.  
```{r}
unlist(str_extract_all(my.email, "[[:digit:]]+"));
```
c. Why expression \\D fails to extract the digits.
#####\\D will match all non-digit occurences.
```{r}
unlist(str_extract_all(my.email, "\\D+"));
```
#####The correct expression is using \\d
```{r}
unlist(str_extract_all(my.email, "\\d+"));
```