---
title: "Violent Crime Rates by US State (1973)"
author: "Arun Kumar"
date: "July 25, 2015"
output: 
  html_document : 
      toc : yes
---

###Objective
To analyze a dataset of US arrest.

###Starting Point
The dataset if obtained from the below site:
[http://vincentarelbundock.github.io/](http://vincentarelbundock.github.io/Rdatasets/datasets.html)  
Dataset reference: [US Arrests](http://vincentarelbundock.github.io/Rdatasets/csv/datasets/USArrests.csv)

###Obtaining the data
1. Download the data in a csv sheet
```{r eval=FALSE}
theURL <- "http://vincentarelbundock.github.io/Rdatasets/csv/datasets/USArrests.csv";
arrest_data <- read.table(file = theURL, header = TRUE, sep = ",");
head(arrest_data);
#State column is named as 'X". Changing the column name to US_State
colnames(arrest_data)[1] <- "US_State";
head(arrest_data);
#Write the data to a CSV
write.table(arrest_data, file = "us_arrest.csv", sep = ",", row.names = FALSE);
#The saved file is uploaded to githib and below is the URL which will be used hereafter.
#https://raw.githubusercontent.com/arunk13/MSDA-Assignments/master/BridgeCourse/Week4/us_arrest.csv 
```

###The dataset
```{r echo=FALSE}
theURL <- "https://raw.githubusercontent.com/arunk13/MSDA-Assignments/master/BridgeCourse/Week4/us_arrest.csv";
arrest_data <- read.table(file = theURL, header = TRUE, sep = ",");
head(arrest_data);
```

###A brief summary of the dataset
```{r}
summary(arrest_data);
```

###Exploratory analysis
```{r}
# g_murder <- ggplot(data = arrest_data_sub, aes(y = Murder, x = US_State, fill = US_State));
# g_murder + geom_bar(stat = "identity", width = 0.2) + guides(fill = FALSE) + xlab("US States") + ylab("Murder(per 100K)") + ggtitle("Crime rate: Murder rate versus states");
# g_murder_UrbanPop <- ggplot(data = arrest_data, aes(y = Murder, x = UrbanPop));
# g_murder_UrbanPop + geom_point(shape = 1) + geom_smooth(method = lm);
# g_rape_UrbanPop <- ggplot(data = arrest_data, aes(y = Rape, x = UrbanPop));
# g_rape_UrbanPop + geom_point(shape = 1) + geom_smooth(method = lm);
# g_assault_UrbanPop <- ggplot(data = arrest_data, aes(y = Assault, x = UrbanPop));
# g_assault_UrbanPop + geom_point(shape = 1) + geom_smooth(method = lm);
```
* Let's find if there is any relationship between variables - Murder, Assault, Rape and percentage of urban population  

Sort the dataset by urban population
```{r}
arrest_data_sub <- arrest_data[order(-arrest_data$UrbanPop),];
```
```{r echo=FALSE}
require(ggplot2);
observation_count <- 7;
```
  
```{r echo=FALSE}
arrest_data_urbanpop <- head(arrest_data_sub, n = observation_count);
g_urbanpop <- ggplot(data = arrest_data_urbanpop, aes(y = UrbanPop, x = US_State, fill = US_State));
g_urbanpop + geom_bar(stat = "identity", width = 0.2) + guides(fill = FALSE) + xlab("US States") + ylab("Urban Population(in %)") + ggtitle("Top 5 US States by percentage of Urban Population (1973)");
```
  
```{r echo=FALSE}
arrest_data_sub <- arrest_data[order(-arrest_data$Murder),];
arrest_data_urbanpop <- head(arrest_data_sub, n = observation_count);
g_urbanpop <- ggplot(data = arrest_data_urbanpop, aes(y = Murder, x = US_State, fill = US_State));
g_urbanpop + geom_bar(stat = "identity", width = 0.2) + guides(fill = FALSE) + xlab("US States") + ylab("Murder Arrest(per 100K residents)") + ggtitle("Top 5 US States by Murder rate (1973)");
```

  
```{r echo=FALSE}
summary(arrest_data$Murder);
ggplot(data = arrest_data, aes(y = Murder, x = factor(0))) + geom_boxplot(size = 0.5) + xlab("") + ylab("Murder Arrest(per 100K residents)") + coord_flip() + ggtitle("Summary of murder arrests in 1973");
```


_Fact: On an average there was `r round(mean(arrest_data$Murder))` muderer arrest per 100,000 residents in 1973_ 
      
```{r echo=FALSE}
arrest_data_sub <- arrest_data[order(-arrest_data$Assault),];
arrest_data_urbanpop <- head(arrest_data_sub, n = observation_count);
g_urbanpop <- ggplot(data = arrest_data_urbanpop, aes(y = Assault, x = US_State, fill = US_State));
g_urbanpop + geom_bar(stat = "identity", width = 0.2) + guides(fill = FALSE) + xlab("US States") + ylab("Assault arrest(per 100K residents)") + ggtitle("Top 5 US States by Assault arrest (1973)");
```

  
```{r echo=FALSE}
arrest_data_sub <- arrest_data[order(-arrest_data$Rape),];
arrest_data_urbanpop <- head(arrest_data_sub, n = observation_count);
g_urbanpop <- ggplot(data = arrest_data_urbanpop, aes(y = Rape, x = US_State, fill = US_State));
g_urbanpop + geom_bar(stat = "identity", width = 0.2) + guides(fill = FALSE) + xlab("US States") + ylab("Rape arrest(per 100K residents)") + ggtitle("Top 5 US States by Rape arrest (1973)");
```


+ Observation
Out of the top 5 states by urban population-
1. California is among one of the top 5 states by rape arrest.

Extending the above study to top 7 states:
1. California is among one of the top 7 states by rape and assault arrest.

With the available data, it can be concluded that percentage of urban population in cities doesnt have any significant impact on the crime rate.
Detailed analysis has to be done on the data of arrests in California to understand if there was an impact of urban poplulatio on the crime rates. 

+ Let's verify the same observation with the help of a scatter plot on the complete dataset.

```{r echo=FALSE}
g_assault_UrbanPop <- ggplot(data = arrest_data, aes(y = Murder, x = UrbanPop));
g_assault_UrbanPop + geom_point(shape = 1) + geom_smooth(method = lm) +  xlab("Urban Population") + ylab("Murder arrest(per 100K residents)") + ggtitle("Murder arrest versus Urban Population (1973)");
```

```{r echo=FALSE}
g_assault_UrbanPop <- ggplot(data = arrest_data, aes(y = Assault, x = UrbanPop));
g_assault_UrbanPop + geom_point(shape = 1) + geom_smooth(method = lm) +  xlab("Urban Population") + ylab("Assault arrest(per 100K residents)") + ggtitle("Assault arrest versus Urban Population (1973)");
```

```{r echo=FALSE}
g_assault_UrbanPop <- ggplot(data = arrest_data, aes(y = Rape, x = UrbanPop));
g_assault_UrbanPop + geom_point(shape = 1) + geom_smooth(method = lm) +  xlab("Urban Population") + ylab("Rape arrest(per 100K residents)") + ggtitle("Rape arrest versus Urban Population (1973)");
```

###Observations
1. The murder and assault arrest does not seem to have any coorelation with urban population.
2. The number of rape arrest seems to have a coorelation with urban pupulation.
3. To strengthen the above findings, simialr study has to be done with latest data.
