---
title: "Standard Deviation"
author: "Arun Kumar"
date: "July 20, 2015"
output: html_document
---
###The question
Card game. Consider the following card game with a well-shuffled deck of cards. If you draw a red card, you win nothing. If you get a spade, you win $5. For any club, you win $10 plus an extra $20 for the ace of clubs.

(a) Create a probability model for the amount you win at this game. Also, find the expected winnings for a single game and the standard deviation of the winnings.

(b) What is the maximum amount you would be willing to pay to play this game? Explain.

###Data frame
```{r echo=FALSE}
X <- c(0,5,10,30);
names(X) <- c("Red", "Spade", "Club", "Spade");
PX <- c(26/52, 13/52, 13/52, 1/52);
names(PX) <- c("P(Red)", "P(Spade)", "P(Club)", "P(Spade)");
random_variables <- data.frame(cbind(X,PX), row.names = NULL);
```
```{r, results='asis'}
knitr::kable(random_variables)
```

###Expected Win
```{r}
expectedWin_X <- round(sum(random_variables$X * random_variables$PX),2);
```
```{r echo=FALSE}
sprintf("The expected win for the above distribution is %.2f", expectedWin_X);
```

###Variance
```{r}
variance_X <- round(sum(random_variables$PX *(random_variables$X - expectedWin_X)^2),2);
```
```{r echo=FALSE}
sprintf("The variance of win for the above distribution is %.2f", variance_X);
```

###Standard deviation
```{r}
standardDeviation_X <- round(sqrt(variance_X),2);
```
```{r echo=FALSE}
sprintf("The standard deviation of win for the above distribution is %.2f", standardDeviation_X);
```