---
title: "Obtain phase: Mushrooms database"
author: "Arun Kumar"
date: "July 19, 2015"
output: 
  html_document: 
    toc: yes
---

###Obtain Munshrooms data from web
Mushrooms data is available from the following link:
https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.data

Reference: [Mushrooms Data Set](https://archive.ics.uci.edu/ml/datasets/Mushroom)

1. Obtain the data from the link:
```{r eval=FALSE}
theURL <- "https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.data;
mushroomsDB <- read.table(theURL, header = FALSE,  sep = ",", stringsAsFactors = FALSE);
write.table(mushroomsDB, file = "Mushroom.csv", sep = ",");
```

2. Upload the data to Github.
3. Use the "raw" url of the CSV file from Github, else the data retrieved will be wrong.

4. Obtain data from Github url.show

```{r}
theURL <- "https://raw.githubusercontent.com/arunk13/MSDA-Assignments/master/BridgeCourse/Week3/Mushroom.csv";
mushroomsDB <- read.table(theURL, header = TRUE,  sep = ",", stringsAsFactors = FALSE);
```

###Overview of the mushroom dataset
Number of rows in the dataset    : `r nrow(mushroomsDB)`  
Number of columns in the dataset : `r ncol(mushroomsDB)`

###Subsetting the dataset
subsetting to a smaller data frame of 100 rows and 7 columns.
The columns chosen for this analysis are:
"class", "cap-shape", "cap-surface", "cap-color", "odor", "population", "habitat"
```{r}
mushroomsDB <- mushroomsDB[1:10, c(1, 2, 3, 4, 6, 22, 23)]
names(mushroomsDB) <- c("class", "cap-shape", "cap-surface", "cap-color", "odor", "population", "habitat");
```
A snapshot of the subset
```{r}
head(mushroomsDB)
```

###Data transformation
To transfor each key in column data to their respective values, the below functions are used.

```{r}
#Below function transforms the class data
transformClassData <- function(key){
  switch (key,
    'p' = 'poisonous',
    'e' = 'edible'
  )
}

#Below function transforms the cap-shape column
transformCapShapeData <- function(key){
  switch (key,
          'b' = 'bell',
          'c' = 'conical',
          'x' = 'convex',
          'f' = 'flat',
          'k' = 'knobbed',
          's' = 'sunken'
  )
}

#Below function transforms the cap-surface column
transformCapSurfaceData <- function(key){
  switch (key,
          'f' = 'fibrous',
          'g' = 'grooves',
          'y' = 'scaly',
          's' = 'smooth'
  )
}

#Below function transforms the cap-color column
transformCapColorData <- function(key){
  switch (key,
          'n' = 'brown',
          'b' = 'buff',
          'c' = 'cinnamon',
          'g' = 'gray',
          'r' = 'green',
          'p' = 'pink',
          'u' = 'purple',
          'e' = 'red',
          'w' = 'white',
          'y' = 'yellow'
  )
}

#Below function transforms the odor column
transformOdorData <- function(key){
  switch (key,
          'a' = 'almond',
          'l' = 'anise',
          'c' = 'creosote',
          'y' = 'fishy',
          'f' = 'foul',
          'm' = 'musty',
          'n' = 'none',
          'p' = 'pungent',
          's' = 'spicy'
  )
}

#Below function transforms the population column
transformPopulationData <- function(key){
  switch (key,
          'a' = 'abundant',
          'c' = 'clustered',
          'n' = 'numerous',
          's' = 'scattered',
          'v' = 'several',
          'y' = 'soletary'
  )
}

#Below function transforms the habitat column
transformHabitatData <- function(key){
  switch (key,
          'g' = 'grasses',
          'l' = 'leaves',
          'm' = 'meadows',
          'p' = 'paths',
          'u' = 'urban',
          'w' = 'waster',
          'd' = 'woods'
  )
}
```

###Applying data transformation on the mushroom dataset
```{r}
mushroomsDB$class <- sapply(mushroomsDB$class, transformClassData);
mushroomsDB$`cap-shape` <- sapply(mushroomsDB$`cap-shape`, transformCapShapeData);
mushroomsDB$`cap-color` <- sapply(mushroomsDB$`cap-color`, transformCapColorData);
mushroomsDB$`cap-surface` <- sapply(mushroomsDB$`cap-surface`, transformCapSurfaceData);
mushroomsDB$odor <- sapply(mushroomsDB$odor, transformOdorData);
mushroomsDB$population <- sapply(mushroomsDB$population, transformPopulationData);
mushroomsDB$habitat <- sapply(mushroomsDB$habitat, transformHabitatData);
```

A snapshot of transformed data:
```{r}
mushroomsDB
```


