---
title: "Reading a chess tournament cross table"
author: "Arun Kumar"
date: "September 22, 2015"
output: 
    html_document:
      toc: yes
---

```{r echo=FALSE}
library(stringr);
library(plyr);
```

###The raw data
```{r echo=FALSE}
tournamentinfo_master <- readLines('https://raw.githubusercontent.com/arunk13/MSDA-Assignments/master/IS607Fall2015/Project1/tournamentinfo.txt');
```
```{r}
head(tournamentinfo_master, 10);
#Reference: http://stackoverflow.com/questions/9068397/import-text-file-as-single-character-string
tournamentinfo <- tournamentinfo_master; #Creating copy of master data, I will be working on this copy.
```
###Data scrubbing
####1.Removing the "---------" lines.  
The dashed lines are not useful for my analysis and hence I will remove it. 
```{r}
tournamentinfo <- str_replace_all(tournamentinfo, "^-+$", "");
tournamentinfo <- tournamentinfo[sapply(tournamentinfo, nchar) > 0]; #Removing the empty vector elements
```
####2.The header has to go.
I know the various data column names and hence keeping the headers in the file is not needed.
```{r}
tournamentinfo <- tournamentinfo[-c(1:2)];
```
Lets see how the data is looking now:
```{r}
head(tournamentinfo, 4);
```
####3.Re-structuring the data
The data of a single player is spread across two vector elements. I will first restructure the data , so that each vector element has the complete details of a single player.
For this:
1. Mark the actual end of line for each player data by appending an indicator.
```{r}
#Function adds a * at the end of every vector element at even index. which is the actual end of line. 
append_separator <- function(x){
  row_index <- which(tournamentinfo == x);
  if(row_index %% 2 == 0){
    x <- paste(x, "*");
  }else{
    x;
  }
}
tournamentinfo <- sapply(tournamentinfo, append_separator);
```
2. Join all the vector elements into a single string.
```{r}
tournamentinfo <- paste(tournamentinfo, collapse = "");
```
3. Split the string into vectors seperated by the indicator.
```{r}
tournamentinfo <- unlist(strsplit(tournamentinfo, split = "\\*"));
```
Lets see how the data is looking now:
```{r}
head(tournamentinfo, 4);
```
###Subsetting the data
1. I do not need the data after post game rating.
```{r}
tournamentinfo <- unlist(str_extract_all(tournamentinfo, "^(.+?\\|){12}"));
```
Lets see how the data is looking now:
```{r}
head(tournamentinfo, 1);
```
###Regular expressions - A powerful tool
Now, as I have the data cleaned up, I can grab the various data elements from the string using regular expression.
```{r}
#Expression to grab the ID.
id <- "^\\s+(\\d+)\\s+\\|"; 
name <- "\\s+([[:alnum:]- ]+)\\b\\s*\\|";
total_points <- "(\\d+\\.?\\d*)\\s*\\|";
#Expression to grab result and opponent number. The data is in the format "|W  39|""
res_opp <- "(\\w)\\s*(\\d{1,})?\\|"; 
state <- "\\s*(\\w{2})\\s*\\|";
#Expression to avoid some part which I dont need.
avoid <- "\\s*.+R:"; 
pre_rating <- "\\s*(\\d{1,}).*->";
post_rating <- "\\s*(\\d{1,}).*\\|";
#Creating the final regex string.
regex_pattern <- paste(id,name,total_points,res_opp,res_opp,res_opp,res_opp,res_opp,res_opp,res_opp,state,avoid,pre_rating,post_rating, sep = "");
#Using str_match , getting the data and putting into a data frame. Using stringsAsFactors is very important, as i faced a lot of trouble using the data in further calculations as the default opponent numbers were in factor format. 
tournamentinfo_final <- as.data.frame(str_match(tournamentinfo, regex_pattern)[,-1], stringsAsFactors = FALSE);
```
###Tidy data  
Let's look at the neat data now.  
```{r}
head(tournamentinfo_final, n = 2);
```
*I am not renaming the columns yet as there are many columns in the data set which I dont need.*    
###Average Pre Chess Rating of Opponents    
The average pre-chess rating of opponents is calculated by adding all the pre-tournament opponents’ ratings and dividing by the total number of games played.
```{r}
#Function to calculate the average pre-chess rating of opponents
calc_avg_pre_opprating <- function(x){
    total_pre_rating <- 0; 
    total_opp <- 0; #Counter for number of games
    opp_columns <- c(5, 7,9,11,13,15,17); #The columns where opponent ID are stored for each player. 
    for(i in opp_columns){
      index <- as.character(x[i][1,1]); #The opponent ID which can be used to find the opponents data row.
      if(!is.na(index)){
        total_pre_rating = total_pre_rating + as.numeric(tournamentinfo_final[index,19]); #The opponent ID is used to get         player's pre-game rating
        total_opp = total_opp + 1;
      }
    }
    round(total_pre_rating/total_opp, 0); # The average
}

tournamentinfo_final <- adply(tournamentinfo_final,.margins = 1, .fun = calc_avg_pre_opprating);
```
###Finally !!!  
I will subset the data frame to include only the rows required.  
```{r}
tournamentinfo_final <- tournamentinfo_final[c(1,17,2,18,20)];
names(tournamentinfo_final) <- c("Player Name", "State", "Rating", "Pre Rating", "Avg Opp Pre-Rating");
head(tournamentinfo_final, 6);
```
Writing the data into csv.
```{r eval=FALSE}
write.csv(tournamentinfo_final, file = "ChessReport.csv");
```