---
title: "Student Performance Data Set"
author: "Arun Kumar"
date: "September 13, 2015"
output: 
  html_document :
    toc : yes

---

###Required R libraries
```{r echo=FALSE}
library(data.table);
library(stringr);
library(ggplot2);
```



###About the dataset

Dataset Name  :   "Student Performance Data Set".  

Source        :   UCI Machine Learning Repository (https://archive.ics.uci.edu/ml/datasets/Student+Performance)  

Information   :   This data approach student achievement in secondary education of two Portuguese schools. The data attributes include student grades, demographic, social and school related features) and it was collected by using school reports and questionnaires. Two datasets are provided regarding the performance in two distinct subjects: Mathematics (mat) and Portuguese language (por). In [Cortez and Silva, 2008], the two datasets were modeled under binary/five-level classification and regression tasks. Important note: the target attribute G3 has a strong correlation with attributes G2 and G1. This occurs because G3 is the final year grade (issued at the 3rd period), while G1 and G2 correspond to the 1st and 2nd period grades. It is more difficult to predict G3 without G2 and G1, but such prediction is much more useful (see paper source for more details). 

*The above descrition is taken from the above mentioned source*

Attribute Information:

*Attributes for both student-mat.csv (Math course) and student-por.csv (Portuguese language course) datasets:* 
1 school - student's school (binary: 'GP' - Gabriel Pereira or 'MS' - Mousinho da Silveira)   
2 sex - student's sex (binary: 'F' - female or 'M' - male)  
3 age - student's age (numeric: from 15 to 22)  
4 address - student's home address type (binary: 'U' - urban or 'R' - rural)  
5 famsize - family size (binary: 'LE3' - less or equal to 3 or 'GT3' - greater than 3)   
6 Pstatus - parent's cohabitation status (binary: 'T' - living together or 'A' - apart)  
7 Medu - mother's education (numeric: 0 - none, 1 - primary education (4th grade), 2 – 5th to 9th grade, 3 – secondary education or 4 – higher education)  
8 Fedu - father's education (numeric: 0 - none, 1 - primary education (4th grade), 2 – 5th to 9th grade, 3 – secondary education or 4 – higher education)  
9 Mjob - mother's job (nominal: 'teacher', 'health' care related, civil 'services' (e.g. administrative or police), 'at_home' or 'other')   
10 Fjob - father's job (nominal: 'teacher', 'health' care related, civil 'services' (e.g. administrative or police), 'at_home' or 'other')   
11 reason - reason to choose this school (nominal: close to 'home', school 'reputation', 'course' preference or 'other')   
12 guardian - student's guardian (nominal: 'mother', 'father' or 'other')   
13 traveltime - home to school travel time (numeric: 1 - <15 min., 2 - 15 to 30 min., 3 - 30 min. to 1 hour, or 4 - >1 hour)   
14 studytime - weekly study time (numeric: 1 - <2 hours, 2 - 2 to 5 hours, 3 - 5 to 10 hours, or 4 - >10 hours)  
15 failures - number of past class failures (numeric: n if 1<=n<3, else 4)   
16 schoolsup - extra educational support (binary: yes or no)   
17 famsup - family educational support (binary: yes or no)   
18 paid - extra paid classes within the course subject (Math or Portuguese) (binary: yes or no)   
19 activities - extra-curricular activities (binary: yes or no)  
20 nursery - attended nursery school (binary: yes or no)   
21 higher - wants to take higher education (binary: yes or no)   
22 internet - Internet access at home (binary: yes or no)   
23 romantic - with a romantic relationship (binary: yes or no)   
24 famrel - quality of family relationships (numeric: from 1 - very bad to 5 - excellent)   
25 freetime - free time after school (numeric: from 1 - very low to 5 - very high)   
26 goout - going out with friends (numeric: from 1 - very low to 5 - very high)   
27 Dalc - workday alcohol consumption (numeric: from 1 - very low to 5 - very high)   
28 Walc - weekend alcohol consumption (numeric: from 1 - very low to 5 - very high)   
29 health - current health status (numeric: from 1 - very bad to 5 - very good)   
30 absences - number of school absences (numeric: from 0 to 93)    

*These grades are related with the course subject, Math or Portuguese:*  
31 G1 - first period grade (numeric: from 0 to 20)   
31 G2 - second period grade (numeric: from 0 to 20)   
32 G3 - final grade (numeric: from 0 to 20, output target)  


###Obtaining the dataset

Both the dataset "student-mat.csv" and "student-por.csv" are downloaded from the source data folder and uploaded to github.  
Below are the Github urls:  
1. https://raw.githubusercontent.com/arunk13/MSDA-Assignments/master/IS607Fall2015/Assignment3/student-mat.csv  
2. https://raw.githubusercontent.com/arunk13/MSDA-Assignments/master/IS607Fall2015/Assignment3/student-por.csv  

Downloading the data from github into R environment  

```{r download-data}
math_student_data_url <- "https://raw.githubusercontent.com/arunk13/MSDA-Assignments/master/IS607Fall2015/Assignment3/student-mat.csv";
math_student_data <- read.table(file = math_student_data_url, header = TRUE, sep = ";");
por_student_data_url <- "https://raw.githubusercontent.com/arunk13/MSDA-Assignments/master/IS607Fall2015/Assignment3/student-por.csv";
por_student_data <- read.table(file = por_student_data_url, header = TRUE, sep = ";");
```


###Knowing our data     
  
#####Size of the dataset
Number of observations    : `r dim(math_student_data)[1]`;  
Number of variables       : `r dim(math_student_data)[2]`;  

#####A snapshot of the data
```{r data-snapshot}
head(math_student_data, n = 3);
str(math_student_data);
```


###Data transformation

#####Subsetting the dataset
For my ease, I will pick few columns which I believe are the most important variables in the dataset.

Variables - #1, #2, #3, #6, #7, #8, #13, #14, #15, #21, #22, #30, #31, #32, #33

```{r data-subset1}
working_student_dataset <- math_student_data[c(1, 2, 3, 6, 7, 8, 13, 14, 15, 21, 22, 30, 31, 32, 33)];
```

We will be working on the new dataset "working_student_dataset".

```{r}
head(working_student_dataset, n = 3);
```

#####Renaming few variables  
I will like to rename the below variables to make it more understandable. Below are the variables and their actual description.  
Pstatus -> parent's cohabitation status.
Medu    -> mother's education
Fedu    -> father's education
G1      -> first period grade
G2      -> second period grade
G3      -> final grade

```{r data-column-rename}
#Using the setnames function in data.table package for easy group renaming using header names instead of column index.
setnames(working_student_dataset, old = c("Pstatus", "Medu", "Fedu", "G1", "G2", "G3"), new = c("parent_cohabitation_status", "mother_edu", "father_edu", "first_period_grade", "second_period_grade", "final_grade"));
```

#####Finding missing values
```{r}
summary(working_student_dataset);
```

*Observation: I dont see any missing values in any of the columns.*

#####Renaming values

The school varibales has two values - GP and MS.  
GP stands for "Gabriel Pereira".  
MS stands for "Mousinho da Silveira" 

Lets use the stringr package for these transformations.

Lets first see what is the class type of school variable.

```{r data-transform-stringr}
class(working_student_dataset$school);
working_student_dataset$school <- str_replace_all(str_c(working_student_dataset$school), c("^GP$" = "Gabriel Pereira", "^MS$" = "Mousinho da Silveira"));
head(working_student_dataset, n = 1);
```

#####Subsetting the data  
Lets subset the data for parent's cohabitaion status - "Apart". I want to see if parent's cohabitation status has any impact on student's grade.

```{r data-subset2}
subset_student_data <- subset(working_student_dataset, working_student_dataset$parent_cohabitation_status == 'A');

#comparing the mean of final grade of overall data with mean of final grade of subsetted data.  
mean(subset_student_data$final_grade) < mean(working_student_dataset$final_grade);
```

###Few analysis

#####Compare mean final grade of two schools
```{r}
aggregate(working_student_dataset[,15], working_student_dataset["school"], mean);
```

#####Compare the final grade distribution of boys versus girls.
```{r echo=FALSE}
student_gender_plot <- ggplot(data = working_student_dataset, aes(x = sex, y = final_grade, fill = sex));
student_gender_plot + geom_boxplot() + xlab("Sex") + ylab("Final Grade") + ggtitle("Distribution of final grade");
```

#####Correlation between grade in the first class and second class.

```{r echo=FALSE}
student_grade_plot <- ggplot(data = working_student_dataset, aes(x = first_period_grade, y = second_period_grade));
student_grade_plot + geom_point() + geom_smooth(method=lm) + xlab("First class grade") + ylab("Second class grade");
```

*Observation: There seems to be a strong correlation between the grades in the first class and second class.*



